Co-moving stars in \textsl{Gaia DR1}

### 2016-10-14 paper/ms.tex
words about projection

For each star with $\snr > 32$, we use the reported parallax and parallax
error to compute a distance corrected for the Lutz-Kelker bias
(\citealt{lutzkelker}):
\begin{equation}
  d_{\rm LK} = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
\end{equation}
We use this distance and the sky positions of each source to compute 3D
Cartesian positions, $\boldsymbol{x}$, for all stars and find the nearest
neighbor to each source in 3-space using the KDTree implementation in
\project{scikit-learn} (\citealt{scikit-learn}). We remove duplicates from
this sample and are left with 29870 pairs of stars. For all pairs, we then
compute the tangential velocity difference
\begin{equation}
  |\Delta \bs{v}_{\rm t}| = d_{\rm LK} \,
                            \sqrt{(\mu_{\alpha,1}-\mu_{\alpha,2})^2 +
                                  (\mu_{\delta,1}-\mu_{\delta,2})^2}
\end{equation}
\figname~\ref{fig:dv-sep} shows the tangential velocity difference against the
physical separation plotted for all pairs that pass the above signal-to-noise
cut. Pairs with small physical separation also tend to have small velocity
differences. Intriguingly, there appear to be an abundance of pairs with small
velocity separation at separations $|\delta\bs{x}| \gtrsim 1~{\rm pc}$.

@@

\todo{More naive visualizations, or save for the full sample?}

>>>>

For each star with $\snr > 32$, we use the reported parallax and parallax
error to compute a distance corrected for the Lutz-Kelker bias
(\citealt{lutzkelker}):
\begin{equation}
  d_{\rm LK} = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
\end{equation}
We use this distance and the sky positions of each source to compute 3D
Cartesian positions, $\boldsymbol{x}$, for all stars and find the nearest
neighbor to each source in 3-space using the KDTree implementation in
\project{scikit-learn} (\citealt{scikit-learn}). We remove duplicates from
this sample and are left with 29870 pairs of stars. For all pairs, we then
compute the tangential velocity difference
\begin{equation}
  |\Delta \bs{v}_{\rm t}| = d_{\rm LK} \,
                            \sqrt{(\mu_{\alpha,1}-\mu_{\alpha,2})^2 +
                                  (\mu_{\delta,1}-\mu_{\delta,2})^2}
\end{equation}
\figname~\ref{fig:dv-sep} shows the tangential velocity difference against the
physical separation plotted for all pairs that pass the above signal-to-noise
cut. Pairs with small physical separation also tend to have small velocity
differences. Intriguingly, there appear to be an abundance of pairs with small
velocity separation at separations $|\delta\bs{x}| \gtrsim 1~{\rm pc}$. Of
course, this figure really only highlights close pairs with small angular
separations:
At large angular separations, the tangential velocity differences will be large
because of projection effects.
To select a more representative population of co-moving stars, we must instead
model the 3-space velocities of the stars given observations of just two
components.

@@
